Modelling the Real Time State of a Public Transport Road Network

### 2017-04-11 realtime-filter/realtime-filter.Rnw
a few wordings

\documentclass{IEEEtran}
%\documentclass[draft,a4paper,onecolumn]{IEEEtran}

@@

\title{Modelling a Public Transport System in Real Time using a Particle Filter}
\author{Tom~Elliott}

@@

\begin{IEEEkeywords}
  Particle filter, transit, real-time
\end{IEEEkeywords}

@@

Basically talk about how the Kalman filter is good, fast, but not as reliable as we want.
Some faults, namely assumptions of normality, symmetry, and that we have to transform observations
(rather than transform parameters to match observations).
Particle filter has been used, and allow us to get around these problems.
Some examples of particle filter usage \ldots

Theory etc of how the particle filter works.

\section{GTFS Data and Road Segmentation}
\label{sec:gtfs}

@@

>>>>

%\documentclass{IEEEtran}
\documentclass[draft,a4paper,onecolumn]{IEEEtran}

@@

\title{Modelling the Real Time State of a Public Transport Road Network}
\author{Tom~Elliott}

@@

\begin{IEEEkeywords}
  Public transport, real-time, particle filter, kalman filter
\end{IEEEkeywords}

@@

Modelling busses in real-time has been done for a while now.
Modelling the underlying state of the road network in order to improve 
travel time predictions has not been explored \ldots
Some looking into complex models for ``future'' travel times \cite{Julio2016}.

The basic idea:
GPS position $\rightarrow$ 
vehicle state (distance, segment, speed) $\rightarrow$
road segment state (speed) $\rightarrow$
travel time predictions

\section{Data Preparation: Road Segmentation}
\label{sec:gtfs}

Real-time data can be used as-is.

However, the shape data needs modification to become composed of \emph{segments},
which are shared between multiple (unrelated) routes.

@@

\bibliographystyle{IEEEtran}
\bibliography{IEEEabrv,../reflist.bib}

### 2017-04-11 realtime-filter/realtime-filter.tex
a few wordings

\documentclass[draft,a4paper,onecolumn]{IEEEtran}\usepackage[]{graphicx}\usepackage[]{color}
%% maxwidth is the original width if it is less than linewidth
%% otherwise use linewidth (to make sure the graphics do not exceed the margin)
\makeatletter
\def\maxwidth{ %
  \ifdim\Gin@nat@width>\linewidth
    \linewidth
  \else
    \Gin@nat@width
  \fi
}
\makeatother

@@

\title{Modelling a Public Transport System in Real Time using a Particle Filter}
\author{Tom~Elliott}
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\begin{document}

@@

Basically talk about how the Kalman filter is good, fast, but not as reliable as we want.
Some faults, namely assumptions of normality, symmetry, and that we have to transform observations
(rather than transform parameters to match observations).
Particle filter has been used, and allow us to get around these problems.
Some examples of particle filter usage \ldots

\section{Particle Filtering}
\label{sec:pf}

Theory etc of how the particle filter works.

\section{GTFS Data and Road Segmentation}
\label{sec:gtfs}

Describe the data (GTFS static and real-time). (Maybe this go in section~\ref{sec:intro}?).

And the idea behind \emph{segmenting} bus routes between intersections.

>>>>

%\documentclass{IEEEtran}
\documentclass[draft,a4paper,onecolumn]{IEEEtran}\usepackage[]{graphicx}\usepackage[]{color}
%% maxwidth is the original width if it is less than linewidth
%% otherwise use linewidth (to make sure the graphics do not exceed the margin)
\makeatletter
\def\maxwidth{ %
  \ifdim\Gin@nat@width>\linewidth
    \linewidth
  \else
    \Gin@nat@width
  \fi
}
\makeatother

@@

\title{Modelling the Real Time State of a Public Transport Road Network}
\author{Tom~Elliott}
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\begin{document}

@@

Modelling busses in real-time has been done for a while now.
Modelling the underlying state of the road network in order to improve 
travel time predictions has not been explored \ldots
Some looking into complex models for ``future'' travel times \cite{Julio2016}.

The basic idea:
GPS position $\rightarrow$ 
vehicle state (distance, segment, speed) $\rightarrow$
road segment state (speed) $\rightarrow$
travel time predictions

\section{Data Preparation: Road Segmentation}
\label{sec:gtfs}

Real-time data can be used as-is.

However, the shape data needs modification to become composed of \emph{segments},
which are shared between multiple (unrelated) routes.

Our work makes use of publically available GTFS data,
which consists of the static component (timetables, shapes, etc),
and the real-time component (vehicle positions and stop time updates).
In its raw form, we are able to obtain the data,
model an individual bus, and determine its (unmeasurable) state
(see section~\ref{sec:transition}).

However, in order to model the network as a whole, 
we need to make some modifications to the data to enable us to 
separate vehicles from routes;
current methodologies, such as those by \ldots,
only use trips belonging to the same route.
In most networks, however,
there is a degree of overlap between routes,
and it is our intention to not only detect these overlaps,
but make full use of them.

We therefore need to be able to break routes into a series
of connecting segments, 
which can then be common between multiple different routes.
The simplest way to find all points at which any two routes 
may converge or diverge is at intersections.
Presently, we are using traffic light positions 
(easily obtained from OSM)
as an approximation for all intersections;
thus, there are some overlapping segments.
